---
title: "CNEP Argentina 2023 — Data Cleaning and Exploration"
subtitle: "Descriptive analysis"
author: "Paloma"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    toc-location: left
    code-fold: show
    code-tools: true
    df-print: paged
    embed-resources: true
    theme: flatly
execute:
  echo: true
  warning: false
  message: false
  cache: true
---

## Setup

```{r}
#| label: setup
#| include: true

library(readxl)
library(dplyr)
library(tidyr)
library(stringr)
library(purrr)
library(here)
library(knitr)
library(DataExplorer)
library(ggplot2)
library(DT)
library(scales)

theme_set(theme_minimal(base_size = 11))
```

# 1. Load raw data

```{r}
#| label: carga

ruta_cnep <- here::here("data", "raw", "Argentina2023_Fin.xlsx")
ruta_dic  <- here::here("data", "raw", "diccionario_cnep_argentina2023.csv")
ruta_list <- here::here("data", "raw", "lista_completa_variables.csv")

cnep   <- read_excel(ruta_cnep)
dic    <- readr::read_csv(ruta_dic,  show_col_types = FALSE)
lista  <- readr::read_csv(ruta_list, show_col_types = FALSE)

cat("CNEP dataset:    ", nrow(cnep),  "rows x", ncol(cnep),  "columns\n")
cat("Dictionary:      ", nrow(dic),   "documented variables\n")
cat("Full var list:   ", nrow(lista), "variables\n")
```

# 2. Data cleaning

## 2.1 Undocumented missing code (90)

Seventeen variables carry an undocumented code `90`, confirmed against the official CNEP v5.0 codebook to not correspond to any substantive response category on any of the seventeen scales.

```{r}
#| label: fix-codigos-na-bloque-a-b

vars_fix_codigo_90 <- c(
  "A.GovPerf_2", "A.EconSit_2", "A.Respsit_2",
  "B.DemSat_2", "B.OneParty_2", "B.TrustMedia_2", "C.LRSelf_2",
  "B.TrustGov_2", "B.PolCompl_2", "B.DontCare_2", "B.DemAuth_2",
  "E.FamTalk_2", "E.FriendTalk_2",
  "J.EqualInd_2", "J.OrderLib_2", "J.PrivPub2_2",
  "B.PolsDefendRich_2"
)

dic <- dic %>%
  mutate(codigos_missing = if_else(
    variable_excel %in% vars_fix_codigo_90 & !str_detect(codigos_missing, "90="),
    paste0(codigos_missing, ", 90=undocumented code"),
    codigos_missing
  ))

readr::write_csv(dic, here::here("data", "raw", "diccionario_cnep_argentina2023.csv"))
cat(sprintf("%d variables corrected in dictionary (90=undocumented code).\n",
    length(vars_fix_codigo_90)))
```

## 2.2 Variable lookup helper

```{r}
#| label: helper-consulta

consultar <- function(variable, diccionario = dic, lista_completa = lista) {
  res <- diccionario |> filter(variable_excel == variable)

  if (nrow(res) > 0) {
    cat("─────────────────────────────────────────────────────────────\n")
    cat("Variable:    ", res$variable_excel, "\n")
    cat("Bloque:      ", res$bloque, "\n")
    cat("Pregunta:    ", res$pregunta, "\n")
    cat("Opciones:    ", res$opciones, "\n")
    cat("Missing:     ", res$codigos_missing, "\n")
    cat("N válidos:   ", res$n_validos, "\n")
    if (!is.na(res$notas) && nchar(res$notas) > 0) {
      cat("Notas:       ", res$notas, "\n")
    }
    cat("─────────────────────────────────────────────────────────────\n")
    return(invisible(res))
  }

  res2 <- lista_completa |> filter(variable == !!variable)
  if (nrow(res2) > 0) {
    cat("─────────────────────────────────────────────────────────────\n")
    cat("Variable:     ", res2$variable, "\n")
    cat("Bloque:       ", res2$bloque, "\n")
    cat("Descripción:  ", res2$descripcion, "\n")
    cat("N válidos:    ", res2$n_validos, "\n")
    cat("(Not documented in detail — inferred description only)\n")
    cat("─────────────────────────────────────────────────────────────\n")
    return(invisible(res2))
  }

  cat("Variable '", variable, "' not found.\n", sep = "")
  invisible(NULL)
}

consultar("B.DontCare_2")
```

## 2.3 Systematic NA recoding

Distinguishes real NA from structural skip codes (`995=skip`, `995=Other`, `998=No pertenece`), which are substantive categories, not missing data.

```{r}
#| label: parser-codigos-na

extraer_codigos_na <- function(texto) {
  if (is.na(texto)) return(numeric(0))
  if (str_detect(texto, regex("^Sin códigos", ignore_case = TRUE))) return(numeric(0))
  if (str_detect(texto, regex("^PENDIENTE", ignore_case = TRUE))) return(numeric(0))
  if (str_detect(texto, regex("^NA si no es", ignore_case = TRUE))) return(numeric(0))

  texto <- str_replace_all(texto, regex("\\d+\\s*=\\s*skip[^,]*", ignore_case = TRUE), "")
  texto <- str_replace_all(texto, regex("\\d+\\s*=\\s*Other", ignore_case = TRUE), "")
  texto <- str_replace_all(texto, regex("\\d+\\s*=\\s*No pertenece[^,]*", ignore_case = TRUE), "")

  str_extract_all(texto, "\\b\\d+\\b")[[1]] |> as.numeric() |> unique()
}

stopifnot(
  identical(extraer_codigos_na("998=Missing, 999=NS"), c(998, 999)),
  identical(extraer_codigos_na("Sin códigos NA"), numeric(0)),
  identical(extraer_codigos_na("995=Other, 998=Missing"), 998),
  identical(extraer_codigos_na("995=skip"), numeric(0)),
  identical(extraer_codigos_na("995=skip, 999=NS"), 999),
  identical(extraer_codigos_na("998=No pertenece, 995=skip"), numeric(0)),
  identical(extraer_codigos_na("PENDIENTE VERIFICAR: 90, 995, 999"), numeric(0)),
  identical(extraer_codigos_na(NA_character_), numeric(0))
)
```

```{r}
#| label: aplicar-recodes-na

cnep_crudo <- cnep

auditoria_na <- tibble(
  variable            = character(),
  codigos_aplicados   = character(),
  casos_recodificados = integer()
)

for (i in seq_len(nrow(dic))) {
  var_name <- dic$variable_excel[i]
  if (!var_name %in% names(cnep)) next

  codigos <- extraer_codigos_na(dic$codigos_missing[i])
  if (length(codigos) == 0) next
  if (!is.numeric(cnep[[var_name]])) next

  mask <- cnep[[var_name]] %in% codigos
  n_rec <- sum(mask, na.rm = TRUE)

  if (n_rec > 0) {
    cnep[[var_name]][mask] <- NA
    auditoria_na <- bind_rows(auditoria_na, tibble(
      variable            = var_name,
      codigos_aplicados   = paste(codigos, collapse = ", "),
      casos_recodificados = n_rec
    ))
  }
}

cat("NAs in raw data:        ", sum(is.na(cnep_crudo)), "\n")
cat("NAs after cleaning:     ", sum(is.na(cnep)), "\n")
cat("New NAs introduced:     ", sum(is.na(cnep)) - sum(is.na(cnep_crudo)), "\n")
cat("Variables affected:     ", nrow(auditoria_na), "\n")
```

```{r}
#| label: auditoria-na

auditoria_na |>
  arrange(desc(casos_recodificados)) |>
  head(30) |>
  knitr::kable(caption = "Top 30 variables recoded to NA")
```

## 2.4 Systematic search for code 90 elsewhere

Documentation only — no recoding applied here.

```{r}
#| label: documentar-codigo-90

extraer_todos_los_numeros <- function(texto) {
  if (is.na(texto)) return(numeric(0))
  if (str_detect(texto, regex("^Sin códigos", ignore_case = TRUE))) return(numeric(0))
  str_extract_all(texto, "\\d+")[[1]] |> as.numeric() |> unique()
}

variables_con_90 <- tibble(
  variable = character(), bloque = character(), pregunta = character(),
  opciones_doc = character(), codigos_missing_doc = character(),
  n_casos_con_90 = integer()
)

for (i in seq_len(nrow(dic))) {
  var_name <- dic$variable_excel[i]
  if (!var_name %in% names(cnep_crudo)) next
  if (!is.numeric(cnep_crudo[[var_name]])) next
  if (var_name == "departamentoId") next

  documentados <- union(
    extraer_todos_los_numeros(dic$codigos_missing[i]),
    extraer_todos_los_numeros(dic$opciones[i])
  )
  if (90 %in% documentados) next

  n_90 <- sum(cnep_crudo[[var_name]] == 90, na.rm = TRUE)
  if (n_90 == 0) next

  variables_con_90 <- bind_rows(variables_con_90, tibble(
    variable = var_name, bloque = dic$bloque[i], pregunta = dic$pregunta[i],
    opciones_doc = dic$opciones[i], codigos_missing_doc = dic$codigos_missing[i],
    n_casos_con_90 = n_90
  ))
}

cat(sprintf("Variables with undocumented code 90: %d\n\n", nrow(variables_con_90)))

variables_con_90 |>
  arrange(bloque, variable) |>
  knitr::kable(caption = "Variables with undocumented code 90")
```

## 2.5 Critical variable checks

```{r}
#| label: verificacion-critica-na

verificar_na <- function(var) {
  if (!var %in% names(cnep)) {
    cat(sprintf("⚠ %s not in dataset.\n", var))
    return(invisible(NULL))
  }
  antes <- sum(is.na(cnep_crudo[[var]]))
  despues <- sum(is.na(cnep[[var]]))
  cat(sprintf("%-40s  NA before: %5d  ->  NA after: %5d  (+%d)\n",
              var, antes, despues, despues - antes))
}

verificar_na("D.Internet_2")
verificar_na("D.CamPaper_2")
verificar_na("D.CamTV_2")
verificar_na("D.CamRadio_2")
verificar_na("E.DiscNumber_2")
verificar_na("H.InfoTest")
verificar_na("L.SubjClassX")
verificar_na("L.WorkStat")
verificar_na("L.WorkStatNew")
verificar_na("I.PartyContPers_2_8")
verificar_na("I.PartyContPers_2_9")
verificar_na("departamentoId")
verificar_na("F.Aso_Noperteneceaningunax")
```

## 2.6 Missing georeference flag

19 respondents share NA across all georeference variables and `departamentoId = 998`.

```{r}
#| label: dummy-sin-georef

cnep <- cnep |> mutate(sin_georreferencia = is.na(region))

cat("Cases without georeference:", sum(cnep$sin_georreferencia), "\n\n")

cnep |>
  filter(sin_georreferencia) |>
  count(provincia) |>
  knitr::kable(caption = "Province of cases without georeference")
```

## 2.7 Ballotage vote code verification

The `_Fin.xlsx` harmonized coding does not match the original questionnaire (`1=Milei, 2=Massa`). Verified empirically against leader thermometers.

```{r}
#| label: verificacion-ballotage

cnep |>
  filter(C.ThermOpp1Leader_2 <= 10, C.ThermGov1Leader_2 <= 10) |>
  group_by(H.VoteWhichRecent2.Z) |>
  summarise(
    n              = n(),
    therm_milei_md = median(C.ThermOpp1Leader_2),
    therm_milei_mn = round(mean(C.ThermOpp1Leader_2), 2),
    therm_massa_md = median(C.ThermGov1Leader_2),
    therm_massa_mn = round(mean(C.ThermGov1Leader_2), 2),
    .groups = "drop"
  ) |>
  knitr::kable(caption = "Ballotage codes crossed with leader thermometers")
```

Code `11` -> Milei voters (therm. Milei ≈ 8.7, therm. Massa ≈ 0.5, n ≈ 1,022). Code `1` -> Massa voters (therm. Massa ≈ 7.1, therm. Milei ≈ 0.5, n ≈ 986). Codes `10`/`12`/`13` -> blank/null/abstention. Code `993` -> don't know.

# 3. Save cleaned dataset

```{r}
#| label: guardar-final
#| eval: true

dir.create(here::here("data", "processed"), recursive = TRUE, showWarnings = FALSE)
saveRDS(cnep, here::here("data", "processed", "cnep_argentina2023_limpio.rds"))
write.csv(auditoria_na, here::here("data", "processed", "auditoria_recodes_na.csv"), row.names = FALSE)

cat("Saved: data/processed/cnep_argentina2023_limpio.rds\n")
cat(sprintf("Final dimensions: %d rows x %d columns\n", nrow(cnep), ncol(cnep)))
```

# 4. Descriptive exploration (cleaned data)

## 4.1 Dataset structure

```{r}
#| label: estructura-general

n_validos <- sapply(cnep, function(x) sum(!is.na(x)))

resumen_general <- tibble(
  Métrica = c("Rows (respondents)", "Total columns", "Populated columns",
              "Empty columns", "Dictionary-documented variables",
              "Inferred-description variables", "Exact duplicate rows"),
  Valor = c(nrow(cnep), ncol(cnep), sum(n_validos > 0), sum(n_validos == 0),
            nrow(dic), sum(lista$en_diccionario == "NO"), sum(duplicated(cnep)))
)

knitr::kable(resumen_general, caption = "Dataset summary")
```

```{r}
#| label: cobertura-bloques

lista |>
  count(bloque, en_diccionario) |>
  pivot_wider(names_from = en_diccionario, values_from = n, values_fill = 0) |>
  rename(Documentadas = SI, `Descripción inferida` = NO) |>
  mutate(Total = Documentadas + `Descripción inferida`) |>
  arrange(desc(Total)) |>
  knitr::kable(caption = "Variables by thematic block")
```

Blocks A-M follow the CNEP comparative framework (A=country situation, B=democracy, C=ideology/party ID, D=media, E=discussion networks, F=associations, H=vote/interest, I=party contacts, J=values, K=electoral integrity, L=demographics, M=Module 6/CSES). Block `z` is Argentine-team administrative variables. Block N (national identity) was not fielded in Argentina.

```{r}
#| label: introduce
#| fig-height: 4

DataExplorer::introduce(cnep)
DataExplorer::plot_intro(cnep, ggtheme = theme_minimal())
```

`Complete Rows = 0%` reflects questionnaire skip logic, not data quality — conditional questions generate structural NAs by design. `Missing Observations` is the informative indicator here, and already reflects the recoding applied in Section 2.

## 4.2 Variable dictionary

```{r}
#| label: diccionario-tabla

DT::datatable(
  dic |> select(variable_excel, bloque, pregunta, opciones,
                codigos_missing, n_validos, notas),
  filter = "top",
  options = list(pageLength = 15, scrollX = TRUE, language = list(url = NULL)),
  rownames = FALSE,
  caption  = "CNEP Argentina 2023 dictionary (filterable)"
)
```

## 4.3 Sampling weights

```{r}
#| label: pesos-resumen

pesos <- c("L.WtWithin", "PONDERADOR_FINAL", "L.WtCombined", "A.A.WtCombined",
           "L.WTACROSS", "pondedad", "pondeduc", "pondregion", "ponajreg",
           "pondfinal", "pondtotal", "pondtotrunc")

pesos_df <- pesos[pesos %in% names(cnep)]
cnep |> select(all_of(pesos_df)) |> summary()
```

```{r}
#| label: pesos-equivalencias

equiv <- tibble(
  Comparación = c("L.WtWithin vs PONDERADOR_FINAL", "L.WtCombined vs A.A.WtCombined",
                  "pondfinal vs PONDERADOR_FINAL", "pondtotal vs PONDERADOR_FINAL"),
  `Máx diferencia abs` = c(
    max(abs(cnep$L.WtWithin   - cnep$PONDERADOR_FINAL),  na.rm = TRUE),
    max(abs(cnep$L.WtCombined - cnep$A.A.WtCombined),    na.rm = TRUE),
    max(abs(cnep$pondfinal    - cnep$PONDERADOR_FINAL),  na.rm = TRUE),
    max(abs(cnep$pondtotal    - cnep$PONDERADOR_FINAL),  na.rm = TRUE)
  )
) |> mutate(`Idénticos` = ifelse(`Máx diferencia abs` < 1e-9, "SÍ", "NO"))

knitr::kable(equiv, caption = "Weight variable equivalence", digits = 4)
```

```{r}
#| label: pesos-distribucion
#| fig-height: 5

cnep |>
  select(L.WtWithin, L.WtCombined) |>
  pivot_longer(everything(), names_to = "Peso", values_to = "Valor") |>
  ggplot(aes(x = Valor, fill = Peso)) +
  geom_histogram(bins = 50, alpha = 0.7) +
  facet_wrap(~ Peso, scales = "free", ncol = 2) +
  labs(title = "Distribution of the two main weights", x = "Weight value", y = "Frequency") +
  theme(legend.position = "none")
```

## 4.4 Demographics

```{r}
#| label: edad

indicadores_edad <- cnep |>
  summarise(
    n_EDAD = sum(!is.na(EDAD)), n_LAge = sum(!is.na(L.Age)),
    discrepancias = sum(EDAD != L.Age, na.rm = TRUE),
    n_menores_18 = sum(EDAD < 18, na.rm = TRUE),
    n_16_17_voto_legal = sum(EDAD >= 16 & EDAD < 18, na.rm = TRUE),
    n_menores_16 = sum(EDAD < 16, na.rm = TRUE)
  ) |>
  pivot_longer(everything(), names_to = "Indicador", values_to = "Valor")

rangos_edad <- tibble(
  Indicador = c("rango_EDAD", "rango_LAge"),
  Valor = c(paste(min(cnep$EDAD, na.rm = TRUE), "-", max(cnep$EDAD, na.rm = TRUE)),
            paste(min(cnep$L.Age, na.rm = TRUE), "-", max(cnep$L.Age, na.rm = TRUE)))
)

bind_rows(indicadores_edad |> mutate(Valor = as.character(Valor)), rangos_edad) |>
  knitr::kable(caption = "Age indicators (EDAD vs L.Age)")
```

```{r}
#| label: edad-grafico
#| fig-height: 4

cnep |>
  ggplot(aes(x = EDAD)) +
  geom_histogram(binwidth = 2, fill = "steelblue", color = "white") +
  geom_vline(xintercept = c(16, 18), linetype = "dashed", color = "darkred") +
  annotate("text", x = 16, y = 200, label = "Legal voting age\n(16)",
           hjust = -0.1, size = 3, color = "darkred") +
  labs(title = "Age distribution (EDAD)",
       subtitle = paste("N =", sum(!is.na(cnep$EDAD)),
                        "| 16-17:", sum(cnep$EDAD %in% 16:17, na.rm = TRUE),
                        "| <16:", sum(cnep$EDAD < 16, na.rm = TRUE)),
       x = "Age", y = "Frequency")
```

`EDAD` (range 14-90) is the real age including 16-17 year-olds with legal voting rights; used for electoral analysis. `L.Age` filters minors and is not used. 23 respondents under 16 are likely capture errors.

```{r}
#| label: genero

cnep |> count(SEXO, genero, GÉNERO_NUMÉRICO, L.Gender) |>
  knitr::kable(caption = "Four gender variable versions crossed")
```

`genero` (text) identifies the 14 respondents self-identifying as "otro". `SEXO` collapses them into "femenino". `GÉNERO_NUMÉRICO` inverts the coding (1=female, 2=male, 3=other). `L.Gender` (CNEP harmonized) leaves the 14 cases as NA. `genero` for descriptive analysis; `GÉNERO_NUMÉRICO` for models.

```{r}
#| label: educacion

cnep |> count(nivelEducativo, NIVEL_EDUC_NUMÉRICO) |> knitr::kable()
```

```{r}
#| label: geografia

cat("\n--- Province (24 categories + CABA) ---\n")
cnep |> count(provincia, sort = TRUE) |> print(n = Inf)

cat("\n--- region (6 cat) vs regiones (7 cat) ---\n")
cnep |>
  count(region, regiones) |>
  pivot_wider(names_from = regiones, values_from = n, values_fill = 0) |>
  knitr::kable()

cat("\n--- tipozona (urban/rural) ---\n")
cnep |> count(tipozona) |> knitr::kable()
```

`regiones` distinguishes CABA (n=302) from Buenos Aires province (n=1,347); `region` merges them. `BUENOS_AIRES`/`CENTRO`/`CUYO`/`NEA`/`NOA`/`PATAGONIA` are sub-regions defined only within their region.

```{r}
#| label: clase

cnep |> count(L.SubjClass) |> knitr::kable()
```

## 4.5 Media and digital intermediation (Block D)

```{r}
#| label: canales-tradicionales

cnep |> count(D.CamPaper_2, name = "n_diarios")  |> knitr::kable()
cnep |> count(D.CamRadio_2, name = "n_radio")    |> knitr::kable()
cnep |> count(D.CamTV_2,    name = "n_television") |> knitr::kable()
```

```{r}
#| label: canales-tradicionales-grafico
#| fig-height: 5

cnep |>
  select(D.CamPaper_2, D.CamRadio_2, D.CamTV_2) |>
  pivot_longer(everything(), names_to = "Canal", values_to = "Días") |>
  filter(!is.na(Días)) |>
  mutate(Canal = recode(Canal, D.CamPaper_2 = "Diarios", D.CamRadio_2 = "Radio",
                        D.CamTV_2 = "Televisión")) |>
  ggplot(aes(x = factor(Días), fill = Canal)) +
  geom_bar() +
  facet_wrap(~ Canal) +
  labs(title = "Days/week following campaign news",
       subtitle = "0 = never, 7 = every day",
       x = "Days/week", y = "Frequency") +
  theme(legend.position = "none")
```

```{r}
#| label: filtro-internet

cnep |>
  count(D.Internet_2) |>
  mutate(pct = round(100 * n / sum(n), 1)) |>
  knitr::kable(caption = "Received/shared campaign info via web/email/social media/SMS?")
```

Respondents answering "No" (~1,444) skip the platform block and go directly to `D.MediaAgree_2`, producing structural missingness (`995`) in `D.Platform_2_*`.

```{r}
#| label: plataformas

plataformas_nombre <- c("Email", "SMS", "Facebook", "Twitter", "WhatsApp",
                        "Instagram", "TikTok", "Otros", "NS")

uso_plataformas <- map_dfr(1:9, function(i) {
  col <- paste0("D.Platform_2_", i)
  vals <- cnep[[col]]
  tibble(plataforma = plataformas_nombre[i], columna = col,
         n_uso = sum(vals == i, na.rm = TRUE), n_no_uso = sum(vals == 0, na.rm = TRUE),
         n_skip_995 = sum(vals == 995, na.rm = TRUE),
         n_otros = sum(!vals %in% c(0, i, 995), na.rm = TRUE))
})

knitr::kable(uso_plataformas, caption = "Platform use during campaign",
             format.args = list(big.mark = "."))
```

```{r}
#| label: plataformas-grafico
#| fig-height: 4

uso_plataformas |>
  filter(plataforma != "NS") |>
  mutate(pct_uso = round(100 * n_uso / nrow(cnep), 1)) |>
  ggplot(aes(x = reorder(plataforma, pct_uso), y = pct_uso)) +
  geom_col(fill = "steelblue") +
  geom_text(aes(label = paste0(pct_uso, "%")), hjust = -0.1, size = 3.5) +
  coord_flip() +
  ylim(0, max(uso_plataformas$n_uso/nrow(cnep)*100) * 1.15) +
  labs(title = "Digital platform use during the 2023 campaign",
       subtitle = paste0("% of N = ", nrow(cnep)), x = NULL, y = "% of respondents")
```

WhatsApp and Instagram lead (~25% each), Facebook close behind (~24%), Twitter fourth (~9.5%), TikTok last (~7%).

```{r}
#| label: frecuencia-online

cnep |>
  filter(D.InterFriends_2 <= 7) |>
  count(D.InterFriends_2) |>
  knitr::kable(caption = "Days/week receiving info from known contacts (digital)")

cnep |>
  filter(D.InterShareFreq_2 <= 7) |>
  count(D.InterShareFreq_2) |>
  knitr::kable(caption = "Days/week posting opinions about candidates")
```

```{r}
#| label: sesgo-digital-lla

cnep |> count(D.InterBias_2) |>
  knitr::kable(caption = "Did digital info favor any party?")

cnep |>
  filter(D.InterBias_2 %in% c(1, 2)) |>
  count(D.InterPart_2_3) |>
  knitr::kable(caption = "Among those perceiving bias: did it favor LLA?")
```

## 4.6 Vote and political interest (Block H)

```{r}
#| label: voto-distribucion

cat("\n=== Turnout Oct 22 ===\n")
cnep |> count(H.Turnout) |> knitr::kable()

cat("\n=== Vote Oct 22 (HARMONIZED — collapses opposition) ===\n")
cnep |> count(H.VoteWhichRecent) |>
  knitr::kable(caption = "1=Massa/UP, 11=opposition (Milei+Bullrich+others), 993=DK/blank")

cat("\n=== Ballotage vote Nov 19 ===\n")
cnep |> count(H.VoteWhichRecent2.Z) |>
  knitr::kable(caption = "See Section 2.7 for verification")
```

```{r}
#| label: info-test
#| fig-height: 4

cnep |> count(H.InfoTest) |>
  knitr::kable(caption = "Correct answers on political knowledge items (0-4)")

cnep |>
  filter(!is.na(H.InfoTest)) |>
  ggplot(aes(x = factor(H.InfoTest))) +
  geom_bar(fill = "steelblue") +
  labs(title = "Political knowledge (0 = none correct, 4 = all correct)",
       x = "Correct answers", y = "Frequency")
```

## 4.7 Party identification and thermometers

```{r}
#| label: party-id

cnep |> count(C.PartyID_2) |> knitr::kable()

cnep |>
  filter(C.PartyID_2 == 1) |>
  count(C.PartyIDWhich_2, sort = TRUE) |>
  knitr::kable(caption = "Party closeness (among those feeling close to a party)")
```

```{r}
#| label: termometros-grafico
#| fig-height: 6

termometros <- c("Massa" = "C.ThermGov1Leader_2", "Milei" = "C.ThermOpp1Leader_2",
                  "CFK" = "C.ThermGov2Leader_2", "Bullrich" = "C.ThermOpp2Leader_2",
                  "Kicillof" = "C.ThermGov3Leader_2", "Macri" = "C.ThermOpp3Leader_2",
                  "Bregman" = "C.ThermOpp6Leader_2", "Villarruel" = "C.ThermOpp8Leader_2")

term_long <- map_dfr(names(termometros), function(lider) {
  tibble(lider = lider, valor = cnep[[termometros[lider]]])
}) |> filter(valor <= 10)

term_long |>
  ggplot(aes(x = valor, fill = lider)) +
  geom_histogram(binwidth = 1, color = "white") +
  facet_wrap(~ lider, ncol = 4) +
  scale_x_continuous(breaks = 0:10) +
  labs(title = "Leader thermometers (0 = rejection, 10 = affinity)",
       x = "Thermometer", y = "Frequency") +
  theme(legend.position = "none")
```

## 4.8 Missingness on RQ2 variables

```{r}
#| label: missingness-rq2
#| fig-height: 6

vars_rq2 <- c(
  "H.VoteWhichRecent2.Z",
  "D.Internet_2",
  "D.Platform_2_3", "D.Platform_2_4", "D.Platform_2_5", "D.Platform_2_6", "D.Platform_2_7",
  "D.InterFriends_2", "D.InterShareFreq_2", "D.InterBias_2", "D.InterPart_2_3",
  "D.CamPaper_2", "D.CamRadio_2", "D.CamTV_2",
  "EDAD", "GÉNERO_NUMÉRICO", "L.Education", "L.SubjClass", "L.Income.Z",
  "C.LRSelf_2", "H.Interest_2", "H.InfoTest", "A.EconSit_2", "A.Respsit_2",
  "regiones", "tipozona", "L.WtWithin"
)

DataExplorer::plot_missing(cnep |> select(any_of(vars_rq2)), ggtheme = theme_minimal())
```

## 4.9 Complete cases

```{r}
#| label: casos-completos

cnep_rq2 <- cnep |> select(any_of(vars_rq2))

n_total     <- nrow(cnep_rq2)
n_completos <- sum(complete.cases(cnep_rq2))
n_perdidos  <- n_total - n_completos

cat("Total N:                          ", n_total, "\n")
cat("N with all RQ2 vars non-missing:  ", n_completos,
    " (", round(100*n_completos/n_total, 1), "%)\n", sep = "")
cat("N lost to missingness:            ", n_perdidos, "\n\n")

n_ballotage <- sum(cnep$H.VoteWhichRecent2.Z %in% c(1, 11), na.rm = TRUE)
n_ballotage_completo <- cnep |>
  filter(H.VoteWhichRecent2.Z %in% c(1, 11)) |>
  select(any_of(setdiff(vars_rq2, "H.VoteWhichRecent2.Z"))) |>
  complete.cases() |> sum()

cat("Milei/Massa ballotage subsample:  ", n_ballotage, "\n")
cat("Complete cases within subsample:  ", n_ballotage_completo, "\n")
```

# 5. Decisions summary

| Decision | Rationale |
|----|----|
| Use `EDAD` (not `L.Age`) | Includes 16-17 year-olds with legal vote |
| Use `genero` or `GÉNERO_NUMÉRICO` (not `SEXO`) | Captures the "otro" category |
| Use `regiones` (not `region`) | Distinguishes CABA from BA province |
| Use `L.WtWithin` (= `PONDERADOR_FINAL`) | Final raking weight |
| Use `L.Income.Z` (not `L.Income`) | Respects questionnaire codes |
| Ballotage as DV for RQ2 | First round collapses opposition |
| Variable-by-variable NA recoding | Generic rule would break cases like `L.WorkStat` (995=Other is substantive) |
| `cnep_crudo` preserved as backup | Enables before/after comparison |
